Statistical paradoxes in coronavirus case-counts: selection bias, measurement error, and the COVID-19 pandemic

### 2020-09-15 write-up/covid-writeup.tex
Update covid-writeup.tex

\begin{table}[!th]
\begin{tabular}{c | c | l l | l | l l }
& & \multicolumn{2}{c}{Total Tests} &  & \multicolumn{2}{c}{Positive Tests} \\
\cline{3-4} \cline{6-7}
& & NonRandom & Random & Census & NonRandom & Random \\ \hline
\multirow{2}{*}{Sex$^\star$} & Female & 57\% & 55\% & 49\% & 53\% & - \\
& Male & 43 & 45 & 51 & 47 & -  \\ \hline
\multirow{3}{*}{Age$^\star$} & $<40$ & 47 & 28 & 53 & 47\\
& $40-59$ & 28 & 36 & 25 & 30\\
& $\geq 60$ & 25 & 36 & 22 & 13\\ \hline
\multirow{2}{*}{Race$^\star$} & White & 60 & 98 & 85 & 50\\
& Nonwhite & 21 & 8 & 15 & 50 \\ \hline
\multirow{2}{*}{Symptoms$^{\star \star}$} & Yes & - & - & - \\
& No & - & - & - \\ \hline
\multirow{2}{*}{Contact$^{\star \star}$} & Yes & 11 & 1 & - \\
& No & 89 & 99 & - \\ \hline
\end{tabular}
\caption{Temp \\
$\star$: ISDH case-count nonrandom sample \\
$\star\star$: May 2--3 nonrandom sample
}
\end{table}

@@

\section{Discussion}
\label{section:discussion}
There is nothing routine about COVID-19, including the corresponding statistical questions.  The goal of this paper was to point out questionable statistical routines.  Precision in reported case-count data gives the illusion of information when what what is needed is quantification of uncertainty. Selection bias leads data analysts to feel certain about incorrect conclusions.  We end with a brief discussion of important related topics.

>>>>

\begin{table}[!th]
\begin{tabular}{c | c | l l | l | l l }
& & \multicolumn{2}{c}{Total Tests} &  & \multicolumn{2}{c}{Positive Tests} \\
\cline{3-4} \cline{6-7}
& & NonRandom & Random & Census & NonRandom & Random \\ \hline
\multirow{2}{*}{Sex$^\star$} & Female & 57\% & 55\% & 49\% & 53\% & - \\
& Male & 43 & 45 & 51 & 47 & -  \\ \hline
\multirow{3}{*}{Age$^\star$} & $<40$ & 47 & 28 & 53 & 47\\
& $40-59$ & 28 & 36 & 25 & 3 0\\
& $\geq 60$ & 25 & 36 & 22 & 13\\ \hline
\multirow{2}{*}{Race$^\star$} & White & 60 & 98 & 85 & 50\\
& Nonwhite & 21 & 8 & 15 & 50 \\ \hline
\multirow{2}{*}{Symptoms$^{\star \star}$} & Yes & 62 & 43 & - & 83 & 62 \\
& No & 38 & 57 & - & 17 & 38  \\ \hline
\multirow{2}{*}{Contact$^{\star \star}$} & Yes & 11 & 1 & - \\
& No & 89 & 99 & - \\ \hline
\end{tabular}
\caption{Temp \\
$\star$: ISDH case-count nonrandom sample \\
$\star\star$: May 2--3 nonrandom sample
}
\end{table}

% Random sample
% Symptom 0, Test 0 2040
% Symptom 0, Test 1 18
% Symptom 1, Test 0 1518
% Symptom 1, Test 1 29
% Total symptom 1; (1518+29)/3605 = 43
% Pos test symptom 1; (29)/(29+18) = 62

% Avg 100 COVID-19 hospitalizations in state in that time
% 650 cases
% percent of cases in nonrandom sample with symptoms = 184/(184+37) = 83%
% Then 83% * 650 = 540 symptomatic cases
% 100/540 = 19% of symptomatic cases were hospitalized

% Example: prior to April 28th, only symptoms and a few others
% 04/11: around 500 cases
% 04/11: 117 hospitalizations -> 615 cases

@@

\section{Discussion}
\label{section:discussion}

{\bf Our goal with surveillance is to identify areas of weakness; but selection bias is an issue!}

There is nothing routine about COVID-19, including the corresponding statistical questions.  The goal of this paper was to point out questionable statistical routines.  Precision in reported case-count data gives the illusion of information when what what is needed is quantification of uncertainty. Selection bias leads data analysts to feel certain about incorrect conclusions.  We end with a brief discussion of important related topics.
